---
title: "Advanced_R_markdown"
author: "Saranya"
date: "November 10, 2018"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(purrr)
```
**map_chr(returns character), map_lgl(returns logical), map_dbl(returns number)**
**first arugument as vector and second arguement as function to be applied on each vector or list elements.**
```{r}
map_chr(c(5,3,1,4,2), function(x){
  c("one","two","three","four","five")[x]
})

map_lgl(c(1,2,3,4,5), function(x){
  x>3
})

map_dbl(c(5,10,25), function(x){
  sqrt(x)
})
```

**map_if() function takes 3 arguments.**
**1 - a list or vector containing data,**
**2 - a predicate function (only when vector element satisfies this condition, 3rd function will be applied.)**
**3 - a function to be applied.**
**map_if() always returns a list..**

```{r}
map_if(c(1:5), function(x){
  x %% 2 ==0
}, function(y){
  y^2
}) %>% unlist()
```

**map_at() function takes 3 arguments**
**1. list or vector**
**2. index locations at which the function should be applied**
**3. function to be applied.**
```{r}
map_at(seq(100,500,100), c(1,5), function(x){
  x-50
}) %>% unlist()
```

**map2_chr() takes 2 vectors as input of equal length and third will be function that should be applied on both**
```{r}
map2_chr(letters, 1:26, paste0) 

map2_chr(c(1:5),c(6:10),sum)

map2_dbl(c(1:5),c(11:15),sum)
```
**pmap is similar to map2 but we can have more number of list input elements. #list args should be list of list.**
```{r}
pmap_chr(list(list(1,2,3),list(10,11,12),list('a','b','c')),paste)
```
**reduce() takes first element of vector and implements the function with second value and result will enter the function with 3rd value and so on.**
**the function should take atleast 2 args and returns a vector or list**
```{r}
reduce(c(1:5), function(x,y){
  #message("X is ",x)
  #message("Y is ",y)
  x+y
})

reduce_right(c(1:5), function(x,y){
  message("X is ",x)
  message("Y is ",y)
  x+y
})
```
**contains() searches for specific element inside a vector -> this is not working**
**has_element() also searches for specific element -> this works**
```{r}
#contains(letters,"a")
#contains(letters, "A")

has_element(c(1,2,3,4,5),5)
```
**detect returns the vector element which first satifies the predicate function**

```{r}
detect(10:20, function(x){
  x>18
})
```

**detect_index returns the vector index which first satifies the predicate function.**
```{r}
detect_index(10:20, function(x){
  x>18
})

detect_index(10:20, function(x){
  x>18
})
```

**Filter function**
**keep(), discard(), every(), some()**

**returns only the elements which satisfied predicate function**

```{r}
keep(c(1:10), function(x){
  x %% 5 ==0
})
```
**returns only the elements which doesn't satify the predicate function**

```{r}
discard(c(1:10), function(x){
  x %%5 ==0
})
```

**returns true only when all elements satifies the condition**
```{r}
every(c(1:10),function(x){
  x %% 1 ==0
})

every(c(1:10),function(x){
  x %% 2 ==0
})
```
**returns true if some elements satisfies the condition.**
```{r}
some(c(1:10),function(x){
  x %% 2 ==0
})
```

**compose() used to combine any number of functions into one function*
```{r}
n_unique <- compose(length, unique)

rep(1:5,1:5)

n_unique(rep(1:5,1:5))
```
**partial() -> you can specify some of the arguments of a function, and then partial() will return a function that only takes the unspecified arguments.**
```{r}
mult_three_numbers <- function(x,y,z){
  x*y*z
}

mult_by_15 <- partial(mult_three_numbers, x=3, y=5)

mult_by_15(z=4)
```
**walk()**
```{r}
walk(c("hi","hello","how are you"),message)
```

**Recursive functions**
**fibonnaci series**
```{r}
fib <- function(n){
  stopifnot(n>0)
  if(n==1){
    0
  } else if(n==2){
    1
  } else{
    fib(n-1)+fib(n-2)
  }
}



fib(1)
fib(2)
fib(3)
fib(4)
fib(5)

map_dbl(1:10,fib)
```
**optimizing fibonnaci series**
```{r}
fib_tbl <- c(0,1,rep(NA,23))

fib_mem <- function(n){
  stopifnot(n>0)
  if(!is.na(fib_tbl[n])){
    fib_tbl[n]
  } else {
    fib_tbl[n-1] <<- fib_mem(n-1)
    fib_tbl[n-2] <<- fib_mem(n-2)
    fib_tbl[n-1] + fib_tbl[n-2]
  }
}

map_dbl(1:10,fib_mem)
```

**comparing fib() and fib_mem() functions**
```{r}
library(purrr)
library(microbenchmark)
library(tidyr)
library(magrittr)
library(dplyr)

fib_data <- map(1:10,function(x){ microbenchmark(fib(x), times =100)$time })
names(fib_data) <- paste0(letters[1:10],1:10)
fib_data <- as.data.frame(fib_data)

fib_data %<>%
  gather(num, time) %>%
  group_by(num) %>%
  summarise(med_time = median(time))


memo_data <- map(1:10, function(x){ microbenchmark(fib_mem(x),times = 100)$time})
names(memo_data) <- paste0(letters[1:10],1:10)
memo_data <- as.data.frame(memo_data)


memo_data %<>% gather(num, time) %>% group_by(num) %>% summarize(med_val = median(time))


plot(1:10, fib_data$med_time, xlab ="Fibonacci number", ylab ="Median time", pch =18, bty="n", xaxt = "n", yaxt ="n")
axis(1, at = 1:10)
axis(2, at = seq(0,3500000, by=50000))
points(1:10+0.1, memo_data$med_val, col ="blue", pch =18)
legend(1, 600000, c("Non Memorized","Memorized"), pch = 15, col=c("black","blue"),bty="n", cex=1, y.intersp = 1.5)
```

**expressions**
**expressions should be mentioned inside quote()**

```{r}
two_plus_two <- quote(2+2)
```
**expressions can be executed using eval**
```{r}
eval(two_plus_two)
```
**parse() to transform string into expression**
```{r}
tpt_string <- "2+2"
tpt_expression <- parse(text = tpt_string)
eval(tpt_expression)
```
**reverse of parse, transforms expression to string**
```{r}
deparse(two_plus_two)
```
**we can modify the contents of expression before evaluation**
```{r}
sum_expr <- quote(sum(10,50))
sum_expr[1]
sum_expr[2]
sum_expr[3]

eval(sum_expr)

sum_expr[[1]] <- quote(paste0)
sum_expr[[2]] <- 10
sum_expr[[3]] <- 50
eval(sum_expr)
```
**we can compose expressions using call()**
```{r}
new_expr <- call("paste0",1,2)
eval(new_expr)
new_expr
```
**match.call() - You can capture the the expression an R user typed into the R console**
**when they executed a function by including match.call() in the function the user executed**
```{r}
return_expr <-  function(...){
  match.call()
}

return_expr(2,col="blue",FALSE) # here while calling the function, whatever parameters are used this will be converted as expression
return_expr(2,col="blue",FALSE)
```
**match.call() is used to capture the expression that the user enters. first argument of the expr is evaluated for if it is numeric**
```{r}
first_arg <- function(...){
  expr <- match.call()
  first_arg_expr <- expr[[2]] # expr[1] corresponds to function name, in this case first_arg
  first_arg <- eval(first_arg_expr)
  if(is.numeric(first_arg)){
    paste("The first argument is", first_arg)
  } else {
    "The first argument is not numeric."
  }
}

first_arg(2,4)
first_arg("sara",5)
```

**creating new environment**
**assign() is used to store values into new environment**
**get() is used to retrieve value from new environment, assign() and get() are opposite**
**we can assign values to environment like handling list**
```{r}
my_personal_env <- new.env()
my_personal_env$a <- 100
get("a", envir =  my_personal_env)

assign("b", 200, envir = my_personal_env)
get("b",envir = my_personal_env)
```
**ls() to list variables of environment**
**exists() to check if a variable is present in environment**
**rm() to remove a variable in environment**
```{r}
ls(envir = my_personal_env)
exists("a",envir = my_personal_env)
rm("b", envir = my_personal_env)
exists("b", envir = my_personal_env)
```

**environments are organized in parent/child relationship, such that each environment keeps track of its parent but parents are unaware of their children**
**search() result indicates, globalenv is child of bindrcpp and bindrcpp is child of dplyr etc**
```{r}
search()

library(ggplot2)
search()
```
**after loading ggplot2, this has become parent of globalenv.

**global assignment operator**
```{r}
a <- 10

my_fun <- function(){
  a<-5
  return(a)
}
my_fun() # returns 5 as the value is from execution environment
a # a value is still 10

my_fun_2 <- function(){
  a <<- 5
}

my_fun_2()
a

#b # no object found
#exists(b)

assign2 <- function(){
  b <<- 25
}
assign2()
#b # returns 25
```
**Error handling**

**stop() to generate an error and stops the execution**
```{r}
#stop("This is an error")
```
**stopifnot evaluates series of expression and halts the execution if any of the expression is false**

```{r}
error_function <- function(x){
  stopifnot(x>0)
  print(x)
}

#error_function(-1)
error_function(5)
```
**warning() - similar to stop() displays message but doesn't halt the execution**
```{r}
warning("This is a warning!")
```
**messages() -> justs prints the text on screen**
```{r}
message("hello!")
```

**tryCatch() block **
**1st arg - expression to evaluate**
**2nd and 3rd args -> error and warning functions**
**finally block which executes by default**
```{r}
my_tryCatch <- function(expr){
  tryCatch(expr, 
           error = function(e){
             message("Error occured, ",e)
           }, 
           warning = function(w){
             message("Warning occured, ",w)
           },
           finally = {
             message("Finally done!")
           }) 
}
my_tryCatch(sum(10,5))

my_tryCatch(sum(5,"sara"))
my_tryCatch(as.numeric(c(5,"sara")))

```